### 2015-02-23 fundamentals/impl.tex
add decay line cache figure and discussion

In large simulations, many material objects may change frequently.  Material
decay can also contribute significantly to such changes.  In order to help
avoid unnecessary runtime performance and database size impacts, compositions
in \Cyclus have some special features.  In particular, compositions are stored
immutably in the database so that multiple material objects can hold references
to the same composition safely.  Additionally, when transmutation and decay
result in new compositions, those are cached for additional avoidance of
redundancy. Composition immutability in concert with decay history caching help
eliminate many redundant decay calculations in addition to reducing the total
number of composition objects recorded in the database.

\TODO{do we want a diagram helping show this composition caching?}

>>>>

In large simulations, many material objects may change frequently.  Material
decay can also contribute significantly to such changes.  In order to help
avoid unnecessary runtime performance and database size impacts, compositions
in \Cyclus have some special features.  In particular, compositions are
immutable once created.  This allows multiple material objects to hold
references to the same composition safely.  Additionally, new compositions
resulting from decay are cached and used to avoid redundant decay
calculations.  Composition immutability in concert with decay history caching
help eliminate many redundant calculations in addition to reducing the
total number of composition entries recorded in the database.

Figure \ref{fig:compositions} shows a set of four compositions A, B, C, and a
yet-uncomputed decayed composition D.  B comes from decaying A 1 time step.  C
comes from decaying B 2 time steps, etc.  Black represents the decay history
cache for this particular composition chain. Blue represents decay
calculations that are satisfied by cache lookups.   If A needs to be decayed 1
time step (A +1), a lookup is performed on the decay line and B is found and
returned - no decay calculation is performed and no new compositions need to
be created or recorded.  Decaying B 3 time steps requires a decay calculation
to compute D, but subsequent requests such as decaying A 4 time steps will not
require any recalculation.

\begin{figure}[htbp!]
\begin{center}
\includegraphics{./images/compositions}
\end{center}
\caption{Composition decay history chain helps avoid redundant decay calculations and reduce duplication in the database.}
\label{fig:composition}
\end{figure}
